### 2020-05-26 intro.tex
mostly refs

\begin{figure*}[tb]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Volume of GPU swaps detected on Titan at individual
    inventories (narrow blue) and yealy sum totals for 2014 and later
    (wide gray). High swap volumes prior to 2014 reflect two major
    rework cycles on mechanical assembly issues at PCIe connectors
    that can be characterized as a break-in period. Our analysis in
    this paper excludes the break-in period and is focused on units in
    place by 2014 or swapped in during the years that follow.} 
  \label{fig:chronology}
\end{figure*}
Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.  The first two rework cycles before 2014
involving blade mechanical assembly can be characterized as a break-in
period on a new system that is first of its kind and pushes many
technological boundaries. The resulting field engineering as well as
data collection changes are responsible for the high GPU swap counts
before 2014. Late 2013 begins a very long period of high reliability
and stable operation with very few issues until about mid 2016, when
GPU failures begin to rise. This results in the final rework cycle,
replacing 11,000 GPUs throughout much of 2017. The peak at the end in
the figure is an artifact of turning off or ``removing'' all the
units as Titan is shut down and disassembled.

>>>>

\begin{figure*}[tb]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{A seven-year tale of GPU life on a leadership class
    supercomputer: Volume of GPU swaps detected on Titan at individual
    inventories (narrow blue) and yealy sum totals for 2014 and later
    (wide gray) over its entire lifetime. High swap volumes prior to
    2014 reflect two major rework cycles on mechanical assembly issues
    at PCIe connectors that can be characterized as a break-in
    period. Our analysis in this paper excludes the break-in period
    and is focused on units in place by the end of 2013 or swapped in
    during the years that follow.}
  \label{fig:chronology}
\end{figure*}
Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories. The first two rework cycles before 2014
involving blade mechanical assembly can be characterized as a break-in
period on a new system that is first of its kind and pushes many
technological boundaries. The resulting field engineering as well as
data collection changes are responsible for the high GPU swap counts
before 2014. Late 2013 begins a very long period of high reliability
and stable operation with very few issues until about mid 2016, when
GPU failures begin to rise. This results in the final rework cycle,
replacing 11,000 GPUs throughout much of 2017. The peak at the end in
the figure is an artifact of turning off or ``removing'' all the
units as Titan is shut down and disassembled.

### 2020-05-26 survival.tex
mostly typos and two graphs
mostly refs

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}
\section{Survival Analysis}
\label{section:survival}
\renewcommand{\pkg}[1]{\textsf{#1}} Survival analysis (SA) methods,
sometimes referred as time to event methods, focus on the effect of
the machine on the GPU units. These methods have roots in biomedical
statistics, where the effect on the patient is most important. So here
we treat the GPUs as patients to determine factors that contribute to
failures. SA methods use and combine information across the
operational lifetimes of all GPUs. For these analyses,
we take apart the location string of each GPU into variables {\em col,
  row, cage, slot, and node} and study the influence of the locations
on the GPU life times. The construction of a GPU life time is more
complex than it initially appears because the units are observed only
at reboot time, because most units were proactively replaced to
prevent failure, and because some units continue in operation after
OTB and DBE events (when a second reboot may be successful).

@@

Kaplan-Meier survival analysis (KM) starts with computing the
probability of survival beyond a given time. It is a nonparametric
technique that makes no specific failure model assumptions, such as
Weibull, Exponential, etc. The technique is able to use censored
observations and can also split the data into subpopulations to
compute separate survival curves.

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
{\em batch}, separate survival curves are computed for each {\em cage}
by {\em node} combination. Along with the survival curve estimate,
this analysis provides 95\% confidence region for survival probability
shaded around the curves.
% Inserting figure here so it appears at top of next page
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em old}
    batch.}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em new}
    batch.}
  \label{fig:cox-new}
\end{figure}

@@

Our interpretation of correlation with details of the cooling system
comes mostly from the hazard ratios for the old batch. The new batch
has mot had many failure events and the uncertainty bars of nearly all
ratios include 1, which is no difference. In the old batch, we see
that {\em cage} has the strongest effect, putting the highest hazard
ratio on cage 2, which is consistent with lowest survivals in the KM
analysis earlier. Its ratio value near 20 has to be interpreted with
caution because it is a time averaged value on a log scale and suffers
from the degrading balance mentioned in the previous paragraph. This
aspect is not captured by the uncertainty, which accounts for the
randomness of the failures and not for the geometric time averaging
\cite{coxhazardinterpret}. Consequently, we interpret the pattern of
relative hazards rather than actual ratio magnitudes. More in-depth
analysis with penalized estimation methods like \cite{bender2019} can
take such balance issues in the exposure history into account and
provide time-dependent hazard estimates.

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}
\section{Survival Analysis}
\label{section:survival}
\renewcommand{\pkg}[1]{\textsf{#1}} Survival analysis (SA) methods
\cite{survival-book}, 
sometimes referred as time to event methods, focus on the effect of
the machine on the GPU units. These methods have roots in biomedical
statistics, where the effect on the patient is most important. So here
we treat the GPUs as patients to determine factors that contribute to
failures. SA methods use and combine information across the
operational lifetimes of all GPUs. For these analyses,
we take apart the location string of each GPU into variables {\em col,
  row, cage, slot, and node} and study the influence of the locations
on the GPU life times. The construction of a GPU life time is more
complex than it initially appears because the units are observed only
at reboot time, because most units were proactively replaced to
prevent failure, and because some units continue in operation after
OTB and DBE events (when a second reboot may be successful).

@@

Kaplan-Meier survival analysis (KM)
\cite{kaplan-meier1958,survival-book} starts with computing the
probability of survival beyond a given time. It is a nonparametric
technique that makes no specific failure model assumptions, such as
Weibull, Exponential, etc. The technique is able to use censored
observations and can also split the data into subpopulations to
compute separate survival curves.

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} \cite{survival-package} and
\pkg{survminer} \cite{survminer-package} for the KM analysis, which is
reported in Figure~\ref{fig:km-all-cage-node}. Within each {\em
  batch}, separate survival curves are computed for each {\em cage} by
{\em node} combination. Along with the survival curve estimate, this
analysis provides 95\% confidence region for survival probability
shaded around the curves.
% Inserting figure here so it appears at top of next page
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em old}
    batch.}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em new}
    batch.}
  \label{fig:cox-new}
\end{figure}

@@

Our interpretation of correlation with details of the cooling system
comes mostly from the hazard ratios for the old batch. The new batch
has not had many failure events and the uncertainty bars of nearly all
ratios include 1, which is no difference. In the old batch, we see
that {\em cage} has the strongest effect, putting the highest hazard
ratio on cage 2, which is consistent with lowest survivals in the KM
analysis earlier. Its ratio value near 20 has to be interpreted with
caution because it is a time averaged value on a log scale and suffers
from the degrading balance mentioned in the previous paragraph. This
aspect is not captured by the uncertainty, which accounts for the
randomness of the failures and not for the geometric time averaging
\cite{coxhazardinterpret}. Consequently, we interpret the pattern of
relative hazards rather than actual ratio magnitudes. More in-depth
analysis with penalized estimation methods like \cite{bender2019} can
take such balance issues in the exposure history into account and
provide time-dependent hazard estimates.
